### 2013-01-31 resources/mmalpha/mmalphaV2-1-0_binary/doc/QuickStart/AlphaStart.tex
Added mmalpha (source and binary)

\documentclass[12pt]{article}
\usepackage{fullpage,epic,eepic,epsfig,makeidx,color,graphicx,moreverb,boxedminipage,version}
\sloppy

@@

\newcommand{\sitemmalfa}
{\texttt{{http://www.irisa.fr/cosi/ALPHA}{http://www.irisa.fr/api/ALPHA}}}
\newcommand{\emailmmalfa}{\texttt{alpha@irisa.fr}}

@@

\title{Getting started with \Alpha{}}
\author{Api, then Cosi, then R2D2 and Compsys\thanks{
Api, Cosi and R2D2 are the names of the 
research groups that successively hosted
research related on \alfa{} and \mmalpha{} at Irisa, Rennes, France. 
From 2001, Compsys in ENS Lyon also participated in 
the development of \mmalpha{}. Recently, Tanguy Risset moved to \textsc{Insa} de Lyon.}
\\Revision of \today{}}
\date{June 2004}
\maketitle

@@

\mmalfa{} is a free software, available under the 
\gpl{}. It can be downloaded from 
the site:

\sitemmalfa{}.

To be run, it requires
\mma{} version 3.0 or later.

@@

\mmalfa{} is still under development at \irisa{} and at ENS Lyon (France).

@@

One important long term goal of \alfa{} is to promote the use of high-level 
functional languages for the synthesis of (parallel) {\vlsi}
architectures. From the short term, Alpha can be useful
for:
\begin{itemize}
\item Providing a correct recurrence equation specification for a particular 
algorithm (see section~\ref{language} and \ref{basic}).
\item Simulating such a specification (see section~\ref{simulation}).\index{Simulation}
\item Transforming and simplifying a recurrence equation specification (see 
sections~\ref{basic} and~\ref{advanced}).
\item Computing on convex polyhedra 
(see also~\texttt{{http://www.irisa.fr/polylib}}).
\index{polylib@\polylib}
\item Scheduling programs and detecting parallelism 
(see section~\ref{schedule}).
\end{itemize}

@@

The remaining of this document is organized as follows. Section~\ref{howwork}
presents the \mmalfa{} interface; section~\ref{install} explains
the installation procedure; in section~\ref{language}, the \alfa{}
language is briefly presented, while in section~\ref{basic}, 
basic operations on \alfa{} programs are described; hardware
generation is presented in section~\ref{hardware}; in section~\ref{advanced},
other transformations of \alfa{} programs are explained, and 
in section~\ref{structured}, we explain how structured \alfa{} programs
can be structured and transformed. Finally, section~\ref{andnow}
concludes and gives some additional references for further 
exploration of \MMAlfa{}.

@@

Specific \C{} functions are used for two purposes: parse and unparse
\alfa{} programs or expressions, and perform computations on
polyhedra. All \C{} functions are called via \mma{}.
Most of them are available separately in the \polylib{}
\index{the polyhedral library@\polylib{}}
polyhedral library which constitutes the computationnal kernel of the
{\mmalfa{}} environment.

\section{Installing \mmalfa{}}
\label{install}
\index{Installation@Installing \mmalfa{}}
Before going on, you
should install the {\mmalfa{}} environment. 
The installation procedure is explained in
the \MMAlfa{} web site (\sitemmalfa{}).

The easiest way to use \mmalfa{} is to access it 
through its notebook interface. 
\index{notebook interface}
To do so, type \texttt{mathematica} under
\texttt{Unix}, or start \mma{}~3.0 in the Programs menu of 
Windows NT\footnote{Any \mma{} version later than 3.0 will work.}.
\footnote{One can also access 
the \mma{} kernel directly. Type \texttt{math} under
\texttt{Unix}, or start \mma{}~3.0~Kernel in the Programs menu of 
Windows NT. An alternative is to use the \mma{} 
kernel via \texttt{emacs}, but then the demonstration notebooks are
not accessible.}

Once \mmalfa{} is installed, open a new \mma{} notebook
and evaluate the command \texttt{start[]} in this notebook\footnote{Type the command, then press simultaneously the Shift and Enter keys.}.
\index{start@\texttt{start[]}}
This command opens a demonstration notebook which is part
of the \mmalfa{} distribution. 
Then open the cell named "Introduction Notebooks"\footnote{For 
the reader who is not yet familiar with \mma{}'s 
notebooks, notebooks are organized as a tree of cells. Each cell is
delimited by a brackets at its right. Selecting a cell (and all 
its subcells) is done by clicking once its bracket. 
To open a cell, double-click the bracket once the cell is selected.}, 
\index{Introduction notebooks}
and click on button "Getting started". This opens a notebook where all 
the examples of this document are shown.
\index{Getting-started notebook}

@@

\subsection*{Systems, variables, domains and parameters}
An {\Alpha} program is a {\em system}. \alfa{} variables are generalized
arrays which can have any shape (not just rectangles).  The set of
indices of the array is called the {\em domain} of the variable.
\index{system}
\index{domain}
\index{variable}
Example~\ref{ex1} below shows the declaration of a variable {\tt a}
whose domain is the set of points $(i,j)$ in the triangle 
$\{0 \leq i \leq j;j \leq 10\}$.
\begin{ex}{~}
\begin{verbatim}
a : {i,j | 0<= i <= j; j <=10} of integer
\end{verbatim}
\label{ex1}
\end{ex} 
In the example of Fig.~\ref{fig1}, domains are indexed
by a parameter \texttt{N} which is declared, together with its
domain, right after the system name. The values of the parameters
can be constrained by any kind of 
affine constraints. In 
general, parameters allow generic descriptions of algorithms to 
be specified.

@@

@@

\paragraph*{Notes}
The allocation is chosen automatically by \mmalpha{}. 
An option allows one to fix the allocation direction. See
the \texttt{projMatrix} option.

@@

>>>>

%postcript file in $MMALPHA/doc/tutorial/QuickStart.ps
\documentclass[12pt]{article}
\usepackage{fullpage,epic,eepic,epsfig,makeidx,color,graphicx,moreverb,boxedminipage,version}
\sloppy

@@

\newcommand{\sitemmalfa}
{\texttt{{http://www.irisa.fr/cosi/ALPHA}}}
\newcommand{\emailmmalfa}{\texttt{alpha@irisa.fr}}

@@

\title{Getting started with \Alpha{}}
\author{Api, then Cosi, then R2D2 and Compsys, then Cairn\thanks{
Api, Cosi, R2D2 and Cairn are the names of the 
research groups that successively hosted
research related on \alfa{} and \mmalpha{} at Irisa, Rennes, France. 
From 2001, Compsys in ENS Lyon also participated in 
the development of \mmalpha{}. Recently, Tanguy Risset moved to \textsc{Insa} de Lyon.}
\\Revision of \today{}}
\date{June 2004}
\maketitle

@@

\mmalfa{} is a free software, available under the 
\gpl{}. 
What is said here applies to Version 2.0 or \mmalfa{}, 
which currently you may get by sending a mail to 
\begin{verbatim}
patrice.quinton@irisa.fr
\end{verbatim}
To be run, it requires
\mma{} version 3.0 or later.

@@

\mmalfa{} is still under development at \irisa{} and at INSA in Lyon (France).

@@

One important long term goal of \alfa{} is to promote the use of high-level 
functional languages for the synthesis of (parallel) {\vlsi}
architectures. From the short term, Alpha can be useful
for:
\begin{itemize}
\item Providing a correct recurrence equation specification for a particular 
algorithm (see Sections~\ref{language} and \ref{basic}).
\item Simulating such a specification (see Section~\ref{simulation}).\index{Simulation}
\item Transforming and simplifying a recurrence equation specification (see 
Sections~\ref{basic} and~\ref{advanced}).
\item Computing on convex polyhedra 
(see also~\texttt{{http://www.irisa.fr/polylib}}).
\index{polylib@\polylib}
\item Scheduling programs and detecting parallelism 
(see Section~\ref{schedule}).
\end{itemize}

@@

The remaining of this document is organized as follows. Section~\ref{howwork}
presents the \mmalfa{} interface; Section~\ref{install} explains
the installation procedure; in Section~\ref{language}, the \alfa{}
language is briefly presented, while in Section~\ref{basic}, 
basic operations on \alfa{} programs are described; hardware
generation is presented in Section~\ref{hardware}; in Section~\ref{advanced},
other transformations of \alfa{} programs are explained, and 
in Section~\ref{structured}, we explain how structured \alfa{} programs
can be structured and transformed. Finally, Section~\ref{andnow}
concludes and gives some additional references for further 
exploration of \MMAlfa{}.

@@

Specific \C{} functions are used for two purposes: parse and unparse
\alfa{} programs or expressions, and perform computations on
polyhedra. All \C{} functions are called via \mma{}.
Most of them are available separately in the \polylib{}
\index{the polyhedral library@\polylib{}}
polyhedral library which constitutes the computationnal kernel of the
{\mmalfa{}} environment. \mma{} also uses the PiP software designed by Paul Feautrier
(See \texttt{http://www.piplib.org/}).

\section{Installing \mmalfa{}}
\label{install}
\index{Installation@Installing \mmalfa{}}
Before going on, you
should install the {\mmalfa{}} environment. 
The installation procedure is explained 
in Appendix~\ref{install} of this document.
\MMAlfa{} can be downloaded from 
the \MMAlfa{} web site (\sitemmalfa{}).

The easiest way to use \mmalfa{} is to access it 
through its notebook interface. 
\index{notebook interface}
To do so, type \texttt{mathematica} under
\texttt{Unix}, or start \mma{} in the Programs menu of 
Windows NT\footnote{Any \mma{} version later than 3.0 will work.}.
We assume in the following that you
use \mma{} through its Notebook interface.

Once \mmalfa{} is installed, open a new \mma{} notebook
and evaluate the command \texttt{start[]} in this notebook\footnote{Type the command, then press simultaneously the Shift and Enter keys.}.
\index{start@\texttt{start[]}}
This command opens a demonstration notebook which is part
of the \mmalfa{} distribution.

Then open the synthesis notebook by clicking on the Synthesis
button (in the Section called {\em New: synthesis notebooks}\footnote{For 
the reader who is not yet familiar with \mma{}'s 
notebooks, notebooks are organized as a tree of cells. Each cell is
delimited by a brackets at its right. Selecting a cell (and all 
its subcells) is done by clicking once its bracket. 
To open a cell, double-click the bracket once the cell is selected.}). In this 
notebook, you will find several demonstrations examples: a FIR filter, 
a distance calculation program, a matrix-vector multiplication, 
a matrix-matrix multiplication, and a string alignment program called
Samba. Currently, the documentation on the \texttt{syn} command is 
available in 
\begin{verbatim}
$MMALPHA/doc/Synthesis/Synthesis.pdf
\end{verbatim}

Another possibility is to open the cell named "Introduction Notebooks", 
\index{Introduction notebooks}
and click on button "Getting started". This opens a notebook where all 
the examples of this document are shown.
\index{Getting-started notebook}

@@

\subsection*{Systems, variables, domains and parameters}
An {\Alpha} program is a {\em system}. \alfa{} variables\footnote{In \Alpha{}, 
identifiers are case sensitive. Undescores are accepted, as in 
\texttt{A\_1} but {\em should be avoided}, as \mma{} identifiers
cannot contain undescores and in some situations, this may result in 
unexpected errors. Moreover, avoid using variables that are distinguished
only by there case, as in \texttt{x} and \texttt{X}; since eventually,
your program will be translated in \vhdl{} that is not case sensitive.} are generalized
arrays which can have any shape (not just rectangles).  The set of
indices of the array is called the {\em domain} of the variable.
\index{system}
\index{domain}
\index{variable}
Example~\ref{ex1} below shows the declaration of a variable {\tt a}
whose domain is the set of points $(i,j)$ in the triangle 
$\{0 \leq i \leq j;j \leq 10\}$.
\begin{ex}{~}
\begin{verbatim}
a : {i,j | 0<= i <= j; j <=10} of integer
\end{verbatim}
\label{ex1}
\end{ex} 
In the example of Fig.~\ref{fig1}, domains are indexed
by a parameter \texttt{N} which is declared, together with its
domain, right after the system name. The values of the parameters
can be constrained by any kind of 
affine constraints. In 
general, parameters allow generic descriptions of algorithms to 
be specified.

@@

A frequent mistake: if you specify in the additional constraints a wrong coefficient, 
the scheduler fails badly with a quite non specific message. For example, say you add
a constraint such as:
\begin{verbatim}
TAD4 == 1
\end{verbatim}
although variable \texttt{A} has only 2 dimensions, then the scheduler fails, since it
does not check the correctness of this constraint and therefore, calls \mma{}'s 
optimizer with wrong arguments.

@@

\paragraph*{Notes}
When \texttt{appSched} is called in its simplest form, the allocation function
is chosen automatically by \mmalpha{}. Sometimes, this allocation is not convenient.
For example, it may happen that the processor dimension is mapped to 
an infinite dimension of the calculation domain.

Some options of \texttt{appSched} allow one to set the allocation direction. 
The \texttt{projMatrix} option allows one to define the projection 
matrix, that is to say, the linear function which defines the indices where
the processor are mapped. For example:
\begin{verbatim}
appSched[ projMatrix -> {{1,0,0}, {0,0,1}} ]
\end{verbatim}
says that the processor indexes will be the first and the third one
(assuming that there is no parameters.)

The \texttt{projVector} option is another way of specifying the 
allocation function:
\begin{verbatim}
appSched[ projVector -> {{0,1,0}} ]
\end{verbatim}
also projects on indexes 1 and 3.

Both \texttt{projVector} and \texttt{projMatrix} 
can be used only when all variables have the same dimension
and are in the same {\em space}.

If necessary, one can apply \texttt{appSched} variable by variable, 
by using the \texttt{variables} option. For example:
\begin{verbatim}
appSched[ variables -> {"A", "b"}, projMatrix -> {{1,0,0}, {0,0,1}} ]
\end{verbatim}
applies the projection only on variables \texttt{A} and \texttt{b}.

Another way of specifying the allocation function is to define
what index has to be replaced by the schedule. For example:
\begin{verbatim}
appSched[ timeDimensions -> {2} ]
\end{verbatim}
says that the second dimension of the variables can be replaced by
the schedule.

This way of specifying the projection can be combined with the \texttt{variables} option:
\begin{verbatim}
appSched[ variables -> { {"A",{2}}, {"b",{1}}, {other,{3}} ]
\end{verbatim}
specifies that for variable \texttt{A}, the second dimension has to be replaced
by the schedulig, for variable \texttt{b}, the first dimension, and for the other
variables, the third dimension.

If none of these possibilities works, the ultimate solution is to do the change of basis by hand. 
Help may be obtained by running 
\begin{verbatim}
appSched[ verbose->True ]
\end{verbatim}
as this command will list the changes of bases that are applied.

\paragraph*{Troubleshooting}
Remember that no change of basis can be done on input or output variables\,! 
Indeed, the change of basis preserves the semantics of an \alfa{} program, and 
this is not compatible with a reindexing of inputs or outputs. If you really want to 
reindex inputs or outputs, just add local variables where you copy the inputs or
store the outputs.

@@

\section{Limitations of \mma{}}
This section contains a few hints to overcome limitations of \mma{} or
to cope with known difficulties. Some of these hints are already mentioned in the
document.

\subsection{The \vhdl{} translater}
Some situations may lead to unexpected errors when using the \texttt{a2v} 
command. Some of them are easy to avoid.

\subsubsection*{Parameters are forbidden!}
First, remember that the main program should not contain parameters anymore. 
The value of the parameters must have been set, before calling \texttt{a2v}, by the \texttt{fixParameter} 
command. (Generating \vhdl{} for parameterized programs is very difficult, except for very simple
cases.)

\subsubsection*{Remove main program\footnote{This is not a joke \ldots}!}
Second, notice that after the \texttt{alpha0ToAlphard} command, the \alfa{} program
is separated into a controller, cells, and a module. All these elements are themselves
called by a main program -- whose name is the name of the initial system -- that cannot always
be translated into \vhdl{}. When calling \texttt{a2v}, the \vhdl{} translator may then fail on 
this main program. This error has no consequence on the generation of \vhdl{} for the other
subsystems. If you want to avoid it, just remove the main program from the library
using the \texttt{removeSystem} command before calling \texttt{a2v}.

\subsubsection*{Put a bound to your calculations!}
\alfa{} allows unbounded programs to be written, filters for example. This happens for example, 
by using variables with domains of the form
\begin{verbatim}
x : { k | 0 <= k } of integer;
\end{verbatim}
Currently, the \texttt{a2v} translator fails when generating a controller for such programs. 
To overcome this, write instead
\begin{verbatim}
x : { k | 0 <= k <= B } of integer;
\end{verbatim}
where \texttt{B} is a parameter that will be set before calling \texttt{a2v}.

\input{../Install/How-to-install}

### 2013-01-31 resources/mmalpha/mmalphaV2-1-0_binary/doc/Tutorial/demos.tex
Added mmalpha (source and binary)

\chapter{The demos of {\mmalfa}}
\label{demos}

{\sc Version of \today{}}

This chapter presents a few demos of
{\mmalfa}.\index{demos,demonstrations} 
There two types of demos: some demos
have a notebook interface, and the others
should be executed in kernel mode. There are also various
documentations available with the \alfa{} archive.

\section{Notebook demos}
All these demonstrations can be accessed directly by 
hyperlinks of the 
so-called {\em master} notebook, which is started by
evaluating the command \texttt{start[]}.

\begin{description}
\item [Introduction.] Introduction to {\Alpha}

\item [Domlib.] Elementary computations on domains with Domlib.

\item [Fir: Design of a FIR filter.]  The FIR demo shows the typical
steps for the synthesis of a linear architecture for a Finite Impulse
Response filter.  The demo comprises the following steps: loading the
initial program, pipelining, scheduling, assign parameter values,
control signal generation, {\alphaz} generation, {\alphard} generation.

\item [matvect: architecture for matrix vector
multiplication.]

The MATVECT demo shows the typical steps for the
synthesis of a linear architecture for a Matrix Vector Multiplication.
The demo comprises the following steps: loading the initial program,
pipelining, scheduling, assign parameter values, control signal
generation,  {\alphaz} generation, {\alphard} generation.

\item[Tutorial\_intro: ]
very introductory tutorial. Corresponds to document~\cite{first-tutorial}.

\item[mma-intro: ]
quick introduction to \mma{}. Explains how Alpha ASTs are represented
in \mma{}. Warning: the abstract syntax presentation may not be as reliable as
the appendix of this manual (see chapter~\ref{appendixalpha}).
\item[Kalman: ]
this example describes an application of \mmalfa{} to a Kalman
filter. This application, quite complex, shows how to use 
structured scheduling. Reference~\cite{parelec} describes the 
background of this application. 
\item[Kalman-Sqrt: ]
this example describes an application of \mmalfa{} to a Kalman
filter, using a so-called square root covariance method. 
This application, quite complex, shows how to use 
structured scheduling. Reference~\cite{cce} describes the 
background of this application.

\item[Neural-Network: ]
this notebook is not ready yet, but will be soon.
\item[Fuzzy-Logic: ]
an application of fuzzy logic to channel equalization. 
\end{description}

Another serie of notebook allows one to become familiar with
different packages of \mmalfa{}: 
\begin{description}
\item[Alpha: ] the Alpha package.
\item[Control: ] the Control package.
\item[Matrix: ] the Matrix package.
\item[ChangeOfBasis: ] the ChangeOfBasis package.
\item[Normalization: ] the Normalization package.
\item[Cut: ] the Cut package.
\item[Decomposition: ] the Decomposition package.
\item[Static: ] the Static package.
\item[PipeControl: ] the PipeControl package. 
\item[Vhdl: ] the Vhdl package.
\item[CheckAlpHard: ] the AlphHard package.
\item[Meta: ] the Meta package.
\item[Schedule: ] the various schedule packages.
\item[dataFlowSchedule: ]

\item[matlib: ]

\item[Visual: ] the Visual package.

@@

\section{kernel demos}
These demos are avaiblable at the following address:
\begin{verbatim}
/udd/alpha/alpha\_beta/Mathematica/demos/demos.html
\end{verbatim}

>>>>

\chapter{The demos of {\mmalfa}}
\label{demos}
This chapter presents a few demos of {\mmalfa}.\index{demos,demonstrations}
There two types of demos, the ones that are to be seen in a bloc note interface and the ones that should be executed in kernel mode. There are also various documentations available with the archive.

\section{Bloc notes demos}

\begin{description}
 \item [Introduction.] Introduction to {\Alpha}
\item [Domlib.] Elementary computations on domains with Domlib.

 \item [FIR: Design of a FIR filter.]  The FIR demo shows the typical
steps for the synthesis of a linear architecture for a Finite Impulse
Response filter.  The demo comprises the following steps: loading the
initial program, pipelining, scheduling, assign parameter values,
control signal generation, {\alphaz} generation, {\alphard} generation.

 \item [MATVECT: architecture for matrix vector
multiplication.] 
The MATVECT demo shows the typical steps for the
synthesis of a linear architecture for a Matrix Vector Multiplication.
The demo comprises the following steps: loading the initial program,
pipelining, scheduling, assign parameter values, control signal
generation,  {\alphaz} generation, {\alphard} generation.

@@

\section{kernel demos}
These demos are avaiblable at the following address:
\htmladdnormallink{file:/udd/alpha/alpha\_beta/Mathematica/demos/demos.html}
{file:/udd/alpha/alpha\_beta/Mathematica/demos/demos.html}

### 2013-01-31 resources/mmalpha/mmalphaV2-1-0_binary/doc/Tutorial/install.tex
Added mmalpha (source and binary)

\chapter{Installing \mmalfa{}}
\label{appendixtechset}
\label{install}

\newcommand{\mmalpha}{{MMalpha}}
\input{install1.tex}

\section{Getting started}
You may use either {\mma} under {\tt emacs}, or directly on the Unix 
system. 
\begin{itemize}
\item under {\tt emacs}, type {\tt  esc-X Mathematica}. 
\item type math (use the {\mma} textual interface).
\item type mathematica (use the {\mma} textual interface).
\end{itemize}
%Do not use a Notebook, as the output of the pretty printer would not
%be available.
When the {\mma} session is started, you can type {\tt goDemo["INIT"]}
for the initiation demonstration.

When the {\mma} session is started, you can type {\tt on[]} or
\texttt{start[]} for the initiation demonstration.

\section{{\mmalpha} file hierarchy}
\includegraphics[width=12cm]{figures/fileHierarchy}

>>>>

\chapter{Installing \mmalfa{}}

\label{appendixtechset}

\label{install}

\newcommand{\mmalpha}{{MMalpha}}

\input{install1.tex}

\section{Getting started}

You may use either {\mma} under {\tt emacs}, or directly on the Unix

system.

\begin{itemize}

\item under {\tt emacs}, type {\tt  esc-X Mathematica}.

\item type math (use the {\mma} textual interface).

\item type mathematica (use the {\mma} textual interface).

\end{itemize}

%Do not use a Notebook, as the output of the pretty printer would not

%be available.

When the {\mma} session is started, you can type {\tt goDemo["INIT"]}

for the initiation demonstration.

When the {\mma} session is started, you can type {\tt on[]} or

\texttt{start[]} for the initiation demonstration.
